Characterizing dynamical transitions by statistical complexity measures based on ordinal pattern transition networks

### 2021-02-16 manuscript_Huang_rev.tex
After submission.

\begin{figure*}
        \centering
        \includegraphics[width=2\columnwidth]{CompEntropyCNormalized_Logistic.pdf}
\caption{\textcolor{red}{(Color online) Normalized complexity -- entropy planes of the three SCMs for the logistic map: (a) $\mathcal{C}_O$, (b) $\mathcal{C}_W$, and (c) $\mathcal{C}_E$. In (c), normalized complexity values larger than 1 indicate numerical artifacts when the control parameter $r$ approaches a bifurcation point, resulting in long transients. The respective embedding dimensions are indicated by different colors (see legend).  \label{fig:CHlogCmax}}}
\end{figure*}

@@

\begin{figure*}
	\centering 
	\includegraphics[width=2\columnwidth]{CompEntropy_fBm.pdf}
\caption{\small{(Color online) Same as in Fig.~\ref{fig:CElogistic}, but for fractional Brownian motion with a Hurst exponent $H \in (0, 1)$ varying with a step size of $\Delta H=0.05$. Note that the normalized entropy is generally decreasing with increasing $H$ (red arrow).}  \label{fig:CEfBm}}
\end{figure*}

\begin{figure*}
	\centering 
	\includegraphics[width=2\columnwidth]{CompEntropy_fGn.pdf}
\caption{\small{(Color online) Same as in Fig.~\ref{fig:CElogistic}, but for fractional Gaussian noise with a Hurst exponent $H \in (0, 1)$ varying with a step size of $\Delta H=0.05$. Note that only a small region of the complexity--entropy planes is covered in this case. }  \label{fig:CEfGn}}
\end{figure*}

@@

\begin{figure*}
        \centering
        \includegraphics[width=2\columnwidth]{CompEntropyCNormalized_fBm.pdf}
\caption{\color{red}(Color online) Same as in Fig.~\ref{fig:CHlogCmax}, but for different realizations of fractional Brownian motion with varying Hurst exponent $H$.  \label{fig:CHfbmCmax}}
\end{figure*}

@@

\begin{figure*}
	\centering 
	\includegraphics[width=2\columnwidth]{rosslerEntropy.pdf}
\caption{\small{Same as Fig. \ref{fig:bifurcation}, but for Poincar\'e sections of the R\"ossler system while varying the control parameter $b$ (see text for details{\color{red}, embedding dimension $D = 6$).} Panel (a) presents a bifurcation diagram based on the $x$ components of all points in the Poincar\'e sections. Panel (b) shows the largest Lyapunov exponent of the system in dependence of $b$.} \label{fig:bifurRossler}}
\end{figure*}

@@

The second set of example time series comprises physiological signals of human electrocardiogram (ECG) recordings collected from patients from \textcolor{red}{the} MIT-BIH Database and the Creighton University Cardiac Center \cite{GoldbergerMITBIH2000}. Heart rate variability suffering from ventricular arrhythmia presents a physiologically highly significant anomaly that is still relatively poorly understood from a dynamical system perspective, showing rich nonlinear properties \cite{smallCSF2002}. {\color{red}The ECG recordings have been obtained at a 250~Hz sampling frequency for about 8.5~minutes \cite{GoldbergerMITBIH2000}. From these long-term recordings, we focus on the three different rhythmic states of normal sinus rhythm (SR), ventricular tachycardia (VT) and ventricular fibrillation (VF) \cite{smallCSF2002}, which have been annotated by skilled cardiologists based on waveform morphology. To reduce possible finite-size effects in estimating SCMs, we study time series of at least $10,000$ time points each. In consequence, out of 18 different patients, we obtain 14 time series showing sinus rhythm (SR) prior to the onset of arrhythmia, 12 records of VT and 17 records of VF. After the annotations, all time series have different lengths, ranging from $14600$ to $159600$ data points in SR, $16000$ to $132000$ in VT, and $14500$ to $124000$ in VF. We do not apply any preprocessing to the data, except for checking the length requirements. We show some segments of time series of each case in Fig. \ref{tseriesECG}. } 
\begin{figure}
	\centering 
	\includegraphics[width=\columnwidth]{TS_ECG.pdf}
\caption{\small{Three representative ECG recording sections (randomly chosen from different rhythmic states): (a) sinus rhythm (SR), (b) VF and VT. The horizontal axis of each panel is the datum number (at 250~Hz \textcolor{red}{sampling rate}) and the vertical axis \textcolor{red}{describes the} ECG surface voltage (in \textcolor{red}{mV}). } \label{tseriesECG}}
\end{figure}

@@

In both sets of time series, the OPTN based SCMs enhance our understanding of the underlying nonlinear system, which can be seen from the results summarized in Figs. (\ref{fig:fluid}, \ref{fig:ecg}). For the fluid data (Fig.~\ref{fig:fluid}), $\mathcal{H}_O$, $\mathcal{H}_W$, $\mathcal{C}_W$, $\mathcal{H}_E$ and $\mathcal{C}_E$ all show consistent changes of the complexity and entropy values between stable wave solution (SW), quasi-periodic 2-frequency amplitude vacillation (AV-2), quasi-periodic 3-frequency amplitude vacillation (AV-3) and chaotic modulated amplitude vacillation (MAV). Specifically, the complexity of these four cases has the following order: SW has the lowest complexity since it shows periodic oscillations with trivial recurrence, while MAV has the highest complexity since it presents chaotically modulated amplitude vacillation. Quasiperiodic states exhibit intermediate complexity values since these two cases have non-trivial recurrences while still featuring a certain degree of regularity. In our case, AV-3 presents a higher complexity than AV-2, which should however be interpreted with care since the AV-2 series originates from an experiment, while the AV-3 series has been taken from a numerical simulation of the same system (hence, there may be a certain level of observational noise present in AV-2 but not in AV-3). The order of complexity values between periodic, quasiperiodic and chaotic solutions still holds for $\mathcal{C}_O$, but the corresponding values become very similar. In particular, it becomes hard for $\mathcal{C}_O$ to identify the difference between AV-2 and AV-3. 
\begin{figure*}
	\centering 
	\includegraphics[width=2\columnwidth]{fluidExample.pdf}
\caption{\color{red}\small{OPTN based entropy and SCM characteristics for experimental fluid data, which are shown for embedding dimensions of $D = 4, 5, 6$. In all subfigures, stable wave (SW, $\blacklozenge$), quasiperiodic 2-frequency amplitude vacillation (AV-2, $\bullet$), quasiperiodic 3-frequency amplitude vacillation (AV-3, $\blacktriangle$) and chaotic modulated amplitude vacillation (MAV, $\blacksquare$) are distinguished. } \label{fig:fluid}}
\end{figure*}

@@

 {\color{red}Meanwhile, we find a mild dependence of the discriminatory power of $\mathcal{C}_O$ and $\mathcal{C}_W$ on the embedding dimension $D$. In particular, in the case of $D=6$ the median values of $\mathcal{C}_O$ of VF are comparable to that of VT (Fig.~\ref{fig:ecg}(b)). On the other hand for $D=4$, the median values of $\mathcal{C}_W$ are all comparable for SR, VF and VT (Fig.~\ref{fig:ecg}(d)). This insufficiency of $\mathcal{C}_O$ and $\mathcal{C}_W$ to detect differences in case of improper embedding dimension may be explained by the fact that $\mathcal{C}_O$ takes only static pattern frequencies into account, while $\mathcal{C}_W$ considers dynamic pattern transitions, which may therefore not be able to properly distinguish subtle differences reflected in the temporal order of patterns. In contrast, $\mathcal{C}_E$ considers both individual pattern frequencies and dynamic pattern transition frequencies. } 
\begin{figure*}
	\centering 
	\includegraphics[width=2\columnwidth]{ecgExample.pdf}
\caption{\color{red}\small{Box plots for OPTN based entropy and SCM characteristics for human ECG recordings, showing well separated medians for each subject group SR, VF and VT. Entropy and SCM scores are shown for embedding dimensions of $D = 4, 5, 6$. } \label{fig:ecg}}
\end{figure*}

@@

\section*{Acknowledgements}
Parts of this work have been financially supported by the National Natural Science Foundation of China (grant no. 11872182, 11835003, 11875132, 61973086), the Natural Science Foundation of Shanghai (grant no. 18ZR1411800). JZ and YZ further acknowledge support by the Shanghai Municipal Science and Technology Major Project (grant no. 2018SHZDZX01) and ZJ Lab. The authors thank J. B. Borges and O. A. Rosso for sharing algorithms for computing $\mathcal{C}_{min}$ and $\mathcal{C}_{max}$\textcolor{red}{, and two anonymous reviewers for helpful comments on the original version of this manuscript}.

@@

\bibliography{ref_Zou,ref_Donner}

>>>>

\begin{figure*}
        \centering
        \includegraphics[width=2\columnwidth]{CompEntropyCNormalized_Logistic.pdf}
\caption{\textcolor{red}{(Color online) Normalized complexity -- entropy planes of the three SCMs for the logistic map: (a) $\mathcal{C}_O^N$, (b) $\mathcal{C}_W^N$, and (c) $\mathcal{C}_E^N$. In (c), normalized complexity values larger than 1 indicate numerical artifacts when the control parameter $r$ approaches a bifurcation point, resulting in long transients. The respective embedding dimensions are indicated by different colors (see legend).  \label{fig:CHlogCmax}}}
\end{figure*}

@@

\begin{figure*}
	\centering 
	\includegraphics[width=2\columnwidth]{CompEntropy_fBm.pdf}
\caption{\small{(Color online) Same as in Fig.~\ref{fig:CElogistic}, but for fractional Brownian motion with a Hurst exponent $H \in (0, 1)$ varying with a step size of $\Delta H=0.05$. Note that the normalized entropy is generally decreasing with increasing $H$ (red arrow). (a, d) $\mathcal{C}_O$ versus $\mathcal{H}_O$, (b, e) $\mathcal{C}_{W}$ versus $\mathcal{H}_{W}$, and (c, f) $\mathcal{C}_{E}$ versus $\mathcal{H}_{E}$. (a-c) for $D = 4, 5$, and (d-f) for $D = 6$. }  \label{fig:CEfBm}}
\end{figure*}

\begin{figure*}
	\centering 
	\includegraphics[width=2\columnwidth]{CompEntropy_fGn.pdf}
\caption{\small{(Color online) Same as in Fig.~\ref{fig:CElogistic}, but for fractional Gaussian noise with a Hurst exponent $H \in (0, 1)$ varying with a step size of $\Delta H=0.05$. Note that only a small region of the complexity--entropy planes is covered in this case. (a, d) $\mathcal{C}_O$ versus $\mathcal{H}_O$, (b, e) $\mathcal{C}_{W}$ versus $\mathcal{H}_{W}$, and (c, f) $\mathcal{C}_{E}$ versus $\mathcal{H}_{E}$. (a-c) for $D = 4, 5$, and (d-f) for $D = 6$. }  \label{fig:CEfGn}}
\end{figure*}

@@

\begin{figure*}
        \centering
        \includegraphics[width=2\columnwidth]{CompEntropyCNormalized_fBm.pdf}
\caption{\color{red}(Color online) Same as in Fig.~\ref{fig:CHlogCmax}, but for different realizations of fractional Brownian motion with varying Hurst exponent $H$. (a) $\mathcal{C}_O^N$, (b) $\mathcal{C}_W^N$, and (c) $\mathcal{C}_E^N$. \label{fig:CHfbmCmax}}
\end{figure*}

@@

\begin{figure*}
	\centering 
	\includegraphics[width=2\columnwidth]{rosslerEntropy.pdf}
\caption{\small{Same as Fig. \ref{fig:bifurcation}, but for Poincar\'e sections of the R\"ossler system while varying the control parameter $b$ (see text for details{\color{red}, embedding dimension $D = 6$).} Panel (a) presents a bifurcation diagram based on the $x$ components of all points in the Poincar\'e sections. Panel (b) shows the largest Lyapunov exponent of the system in dependence of $b$. (c) $\mathcal{H}_O$ and (d) $\mathcal{C}_O$; (e) $\mathcal{H}_W$ and (f) $\mathcal{C}_W$; (g) $\mathcal{H}_E$ and (h) $\mathcal{C}_E$. } \label{fig:bifurRossler}}
\end{figure*}

@@

The second set of example time series comprises physiological signals of human electrocardiogram (ECG) recordings collected from patients from \textcolor{red}{the} MIT-BIH Database and the Creighton University Cardiac Center \cite{GoldbergerMITBIH2000}. Heart rate variability suffering from ventricular arrhythmia presents a physiologically highly significant anomaly that is still relatively poorly understood from a dynamical system perspective, showing rich nonlinear properties \cite{smallCSF2002}. {\color{red}The ECG recordings have been obtained at a 250~Hz sampling frequency for about 8.5~minutes \cite{GoldbergerMITBIH2000}. From these long-term recordings, we focus on the three different rhythmic states of normal sinus rhythm (SR), ventricular tachycardia (VT) and ventricular fibrillation (VF) \cite{smallCSF2002}, which have been annotated by skilled cardiologists based on waveform morphology. To reduce possible finite-size effects in estimating SCMs, we study time series of at least $10,000$ time points each. In consequence, out of 18 different patients, we obtain 14 time series showing sinus rhythm (SR) prior to the onset of arrhythmia, 12 records of VT and 17 records of VF. After the annotations, all time series have different lengths, ranging from $14600$ to $159600$ data points in SR, $16000$ to $132000$ in VT, and $14500$ to $124000$ in VF. We do not apply any preprocessing to the data, except for checking the length requirements. We show some segments of time series of each case in Fig. \ref{tseriesECG}. } 
\begin{figure}
	\centering 
	\includegraphics[width=\columnwidth]{TS_ECG.pdf}
\caption{\small{Three representative ECG recording sections (randomly chosen from different rhythmic states): (a) sinus rhythm (SR), (b) ventricular fibrillation (VF) and (c) ventricular tachycardia (VT). The horizontal axis of each panel is the datum number (at 250~Hz \textcolor{red}{sampling rate}) and the vertical axis \textcolor{red}{describes the} ECG surface voltage (in \textcolor{red}{mV}). } \label{tseriesECG}}
\end{figure}

@@

In both sets of time series, the OPTN based SCMs enhance our understanding of the underlying nonlinear system, which can be seen from the results summarized in Figs. (\ref{fig:fluid}, \ref{fig:ecg}). For the fluid data (Fig.~\ref{fig:fluid}), $\mathcal{H}_O$, $\mathcal{H}_W$, $\mathcal{C}_W$, $\mathcal{H}_E$ and $\mathcal{C}_E$ all show consistent changes of the complexity and entropy values between stable wave solution (SW), quasi-periodic 2-frequency amplitude vacillation (AV-2), quasi-periodic 3-frequency amplitude vacillation (AV-3) and chaotic modulated amplitude vacillation (MAV). Specifically, the complexity of these four cases has the following order: SW has the lowest complexity since it shows periodic oscillations with trivial recurrence, while MAV has the highest complexity since it presents chaotically modulated amplitude vacillation. Quasiperiodic states exhibit intermediate complexity values since these two cases have non-trivial recurrences while still featuring a certain degree of regularity. In our case, AV-3 presents a higher complexity than AV-2, which should however be interpreted with care since the AV-2 series originates from an experiment, while the AV-3 series has been taken from a numerical simulation of the same system (hence, there may be a certain level of observational noise present in AV-2 but not in AV-3). The order of complexity values between periodic, quasiperiodic and chaotic solutions generally still holds for $\mathcal{C}_O$, but the corresponding values become very similar. In particular, it becomes hard for $\mathcal{C}_O$ to identify the difference between AV-2 and AV-3. 
\begin{figure*}
	\centering 
	\includegraphics[width=2\columnwidth]{fluidExample.pdf}
\caption{\color{red}\small{OPTN based entropy and SCM characteristics for experimental fluid data, which are shown for different embedding dimensions, (a) $D = 4$, (b) $D = 5$, and (c) $D = 6$. In all subfigures, stable wave (SW, $\blacklozenge$), quasiperiodic 2-frequency amplitude vacillation (AV-2, $\bullet$), quasiperiodic 3-frequency amplitude vacillation (AV-3, $\blacktriangle$) and chaotic modulated amplitude vacillation (MAV, $\blacksquare$) are distinguished. } \label{fig:fluid}}
\end{figure*}

@@

 {\color{red}Meanwhile, we find a mild dependence of the discriminatory power of $\mathcal{C}_O$ and $\mathcal{C}_W$ on the embedding dimension $D$. In particular, in the case of $D=6$ the median values of $\mathcal{C}_O$ of VF are comparable to that of VT (Fig.~\ref{fig:ecg}(b)). On the other hand for $D=4$, the median values of $\mathcal{C}_W$ are all comparable for SR, VF and VT (Fig.~\ref{fig:ecg}(d)). This insufficiency of $\mathcal{C}_O$ and $\mathcal{C}_W$ to detect differences in case of improper embedding dimension may be explained by the fact that $\mathcal{C}_O$ takes only static pattern frequencies into account, while $\mathcal{C}_W$ considers dynamic pattern transitions, which may therefore not be able to properly distinguish subtle differences reflected in the temporal order of patterns. In contrast, $\mathcal{C}_E$ considers both individual pattern frequencies and dynamic pattern transition frequencies. } 
\begin{figure*}
	\centering 
	\includegraphics[width=2\columnwidth]{ecgExample.pdf}
\caption{\color{red}\small{Box plots for OPTN based entropy and SCM characteristics for human ECG recordings, showing well separated medians for each subject group SR, VF and VT. Entropy and SCM scores are shown for embedding dimensions of $D = 4, 5, 6$. (a) $\mathcal{H}_O$ and (b) $\mathcal{C}_O$; (c) $\mathcal{H}_W$ and (d) $\mathcal{C}_W$; (e) $\mathcal{H}_E$ and (f) $\mathcal{C}_E$. } \label{fig:ecg}}
\end{figure*}

@@

\section*{Supplementary Material}
See Supplementary material: Sec.~SM1 with Figs.~S1 to S3 for the complexity--entropy planes when the self-loops are included in the transition matrix; Sec.~SM2 with Figs.~S4 and S5 for transition curves for the logistic map and the chaotic R\"ossler system when embedding parameters are $D=6$ and $\tau=1$; Sec.~ SM3 with Fig.~S6 for the linear increase of the maximum complexity $\mathcal{C}_{max}$ versus entropy curve for sufficiently large embedding dimension. However, this linear correlation in general is not expected by fractional Brownian motion and fractional Gaussian noise Sec.~ SM3 with Figs.~S7 and S8.

\section*{Acknowledgements}
Parts of this work have been financially supported by the National Natural Science Foundation of China (grant no. 11872182, 11972288, 11835003, 11875132, 61973086), the Natural Science Foundation of Shanghai (grant no. 18ZR1411800). JZ and YZ further acknowledge support by the Shanghai Municipal Science and Technology Major Project and ZJ Lab (grant no. 2018SHZDZX01). The authors thank J. B. Borges and O. A. Rosso for sharing algorithms for computing $\mathcal{C}_{min}$ and $\mathcal{C}_{max}$\textcolor{red}{, and two anonymous reviewers for helpful comments on the original version of this manuscript}.

@@

% \bibliography{ref_Zou,ref_Donner}
%merlin.mbs aipnum4-1.bst 2010-07-25 4.21a (PWD, AO, DPC) hacked
%Control: key (0)
%Control: author (8) initials jnrlst
%Control: editor formatted (1) identically to author
%Control: production of article title (0) allowed
%Control: page (1) range
%Control: year (1) truncated
%Control: production of eprint (0) enabled
\begin{thebibliography}{43}%
\makeatletter
\providecommand \@ifxundefined [1]{%
 \@ifx{#1\undefined}
}%
\providecommand \@ifnum [1]{%
 \ifnum #1\expandafter \@firstoftwo
 \else \expandafter \@secondoftwo
 \fi
}%
\providecommand \@ifx [1]{%
 \ifx #1\expandafter \@firstoftwo
 \else \expandafter \@secondoftwo
 \fi
}%
\providecommand \natexlab [1]{#1}%
\providecommand \enquote  [1]{``#1''}%
\providecommand \bibnamefont  [1]{#1}%
\providecommand \bibfnamefont [1]{#1}%
\providecommand \citenamefont [1]{#1}%
\providecommand \href@noop [0]{\@secondoftwo}%
\providecommand \href [0]{\begingroup \@sanitize@url \@href}%
\providecommand \@href[1]{\@@startlink{#1}\@@href}%
\providecommand \@@href[1]{\endgroup#1\@@endlink}%
\providecommand \@sanitize@url [0]{\catcode `\\12\catcode `\$12\catcode
  `\&12\catcode `\#12\catcode `\^12\catcode `\_12\catcode `\%12\relax}%
\providecommand \@@startlink[1]{}%
\providecommand \@@endlink[0]{}%
\providecommand \url  [0]{\begingroup\@sanitize@url \@url }%
\providecommand \@url [1]{\endgroup\@href {#1}{\urlprefix }}%
\providecommand \urlprefix  [0]{URL }%
\providecommand \Eprint [0]{\href }%
\providecommand \doibase [0]{http://dx.doi.org/}%
\providecommand \selectlanguage [0]{\@gobble}%
\providecommand \bibinfo  [0]{\@secondoftwo}%
\providecommand \bibfield  [0]{\@secondoftwo}%
\providecommand \translation [1]{[#1]}%
\providecommand \BibitemOpen [0]{}%
\providecommand \bibitemStop [0]{}%
\providecommand \bibitemNoStop [0]{.\EOS\space}%
\providecommand \EOS [0]{\spacefactor3000\relax}%
\providecommand \BibitemShut  [1]{\csname bibitem#1\endcsname}%
\let\auto@bib@innerbib\@empty
%</preamble>
\bibitem [{\citenamefont {Zou}\ \emph {et~al.}(2019)\citenamefont {Zou},
  \citenamefont {Donner}, \citenamefont {Marwan}, \citenamefont {Donges},\ and\
  \citenamefont {Kurths}}]{ZouPR2018}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont
  {Zou}}, \bibinfo {author} {\bibfnamefont {R.~V.}\ \bibnamefont {Donner}},
  \bibinfo {author} {\bibfnamefont {N.}~\bibnamefont {Marwan}}, \bibinfo
  {author} {\bibfnamefont {J.~F.}\ \bibnamefont {Donges}}, \ and\ \bibinfo
  {author} {\bibfnamefont {J.}~\bibnamefont {Kurths}},\ }\bibfield  {title}
  {\enquote {\bibinfo {title} {Complex network approaches to nonlinear time
  series analysis},}\ }\href {\doibase 10.1016/j.physrep.2018.10.005}
  {\bibfield  {journal} {\bibinfo  {journal} {Physics Reports}\ }\textbf
  {\bibinfo {volume} {787}},\ \bibinfo {pages} {1--97} (\bibinfo {year}
  {2019})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Kantz}\ and\ \citenamefont
  {Schreiber}(2004)}]{Kantz97}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {H.}~\bibnamefont
  {Kantz}}\ and\ \bibinfo {author} {\bibfnamefont {T.}~\bibnamefont
  {Schreiber}},\ }\href@noop {} {\emph {\bibinfo {title} {Nonlinear Time Series
  Analysis}}},\ \bibinfo {edition} {2nd}\ ed.\ (\bibinfo  {publisher}
  {Cambridge University Press, Cambridge},\ \bibinfo {year} {2004})\BibitemShut
  {NoStop}%
\bibitem [{\citenamefont {Sprott}(2003)}]{Sprott2003}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.~C.}\ \bibnamefont
  {Sprott}},\ }\href@noop {} {\emph {\bibinfo {title} {Chaos and Time-Series
  Analysis}}}\ (\bibinfo  {publisher} {Oxford University Press, Oxford},\
  \bibinfo {year} {2003})\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Zhang}\ and\ \citenamefont
  {Small}(2006)}]{ZhangPRL2006}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.}~\bibnamefont
  {Zhang}}\ and\ \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Small}},\
  }\bibfield  {title} {\enquote {\bibinfo {title} {{Complex network from
  pseudoperiodic time series: Topology versus dynamics}},}\ }\href {\doibase
  10.1103/PhysRevLett.96.238701} {\bibfield  {journal} {\bibinfo  {journal}
  {{Phys. Rev. Lett.}}\ }\textbf {\bibinfo {volume} {{96}}},\ \bibinfo {pages}
  {238701} (\bibinfo {year} {{2006}})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Marwan}\ \emph {et~al.}(2009)\citenamefont {Marwan},
  \citenamefont {Donges}, \citenamefont {Zou}, \citenamefont {Donner},\ and\
  \citenamefont {Kurths}}]{MarwanPLA2009}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {N.}~\bibnamefont
  {Marwan}}, \bibinfo {author} {\bibfnamefont {J.~F.}\ \bibnamefont {Donges}},
  \bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont {Zou}}, \bibinfo {author}
  {\bibfnamefont {R.~V.}\ \bibnamefont {Donner}}, \ and\ \bibinfo {author}
  {\bibfnamefont {J.}~\bibnamefont {Kurths}},\ }\bibfield  {title} {\enquote
  {\bibinfo {title} {Complex network approach for recurrence analysis of time
  series},}\ }\href {\doibase https://doi.org/10.1016/j.physleta.2009.09.042}
  {\bibfield  {journal} {\bibinfo  {journal} {Phys. Lett. A}\ }\textbf
  {\bibinfo {volume} {373}},\ \bibinfo {pages} {4246 -- 4254} (\bibinfo {year}
  {2009})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Donner}\ \emph {et~al.}(2010)\citenamefont {Donner},
  \citenamefont {Zou}, \citenamefont {Donges}, \citenamefont {Marwan},\ and\
  \citenamefont {Kurths}}]{Donner2010a}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {R.~V.}\ \bibnamefont
  {Donner}}, \bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont {Zou}},
  \bibinfo {author} {\bibfnamefont {J.~F.}\ \bibnamefont {Donges}}, \bibinfo
  {author} {\bibfnamefont {N.}~\bibnamefont {Marwan}}, \ and\ \bibinfo {author}
  {\bibfnamefont {J.}~\bibnamefont {Kurths}},\ }\bibfield  {title} {\enquote
  {\bibinfo {title} {{Recurrence networks --- a novel paradigm for nonlinear
  time series analysis}},}\ }\href {\doibase 10.1088/1367-2630/12/3/033025}
  {\bibfield  {journal} {\bibinfo  {journal} {New J. Phys.}\ }\textbf {\bibinfo
  {volume} {12}},\ \bibinfo {pages} {033025} (\bibinfo {year}
  {2010})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Lacasa}\ \emph {et~al.}(2008)\citenamefont {Lacasa},
  \citenamefont {Luque}, \citenamefont {Ballesteros}, \citenamefont {Luque},\
  and\ \citenamefont {Nuno}}]{Lacasa2008}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {L.}~\bibnamefont
  {Lacasa}}, \bibinfo {author} {\bibfnamefont {B.}~\bibnamefont {Luque}},
  \bibinfo {author} {\bibfnamefont {F.}~\bibnamefont {Ballesteros}}, \bibinfo
  {author} {\bibfnamefont {J.}~\bibnamefont {Luque}}, \ and\ \bibinfo {author}
  {\bibfnamefont {J.~C.}\ \bibnamefont {Nuno}},\ }\bibfield  {title} {\enquote
  {\bibinfo {title} {{From time series to complex networks: The visibility
  graph}},}\ }\href {\doibase 10.1073/pnas.0709247105} {\bibfield  {journal}
  {\bibinfo  {journal} {Proc. Natl. Acad. Sci.}\ }\textbf {\bibinfo {volume}
  {105}},\ \bibinfo {pages} {4972--4975} (\bibinfo {year} {2008})}\BibitemShut
  {NoStop}%
\bibitem [{\citenamefont {Nu{\~{n}}ez}\ \emph {et~al.}(2012)\citenamefont
  {Nu{\~{n}}ez}, \citenamefont {Lacasa}, \citenamefont {Patricio},\ and\
  \citenamefont {Luque}}]{Nunez2012}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
  {Nu{\~{n}}ez}}, \bibinfo {author} {\bibfnamefont {L.}~\bibnamefont {Lacasa}},
  \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont {Patricio}}, \ and\
  \bibinfo {author} {\bibfnamefont {B.}~\bibnamefont {Luque}},\ }\bibfield
  {title} {\enquote {\bibinfo {title} {{Visibility Algorithms: A Short
  Review}},}\ }in\ \href {\doibase 10.5772/34810} {\emph {\bibinfo {booktitle}
  {New Front. Graph Theory}}}\ (\bibinfo  {publisher} {InTech},\ \bibinfo
  {year} {2012})\ Chap.~\bibinfo {chapter} {6}, pp.\ \bibinfo {pages}
  {119--152}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Nicolis}, \citenamefont {Cantu},\ and\ \citenamefont
  {Nicolis}(2005)}]{Nicolis2005}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {G.}~\bibnamefont
  {Nicolis}}, \bibinfo {author} {\bibfnamefont {A.~G.}\ \bibnamefont {Cantu}},
  \ and\ \bibinfo {author} {\bibfnamefont {C.}~\bibnamefont {Nicolis}},\
  }\bibfield  {title} {\enquote {\bibinfo {title} {{Dynamical aspects of
  interaction networks}},}\ }\href {\doibase 10.1142/S0218127405014167}
  {\bibfield  {journal} {\bibinfo  {journal} {Int. J. Bifurc. Chaos}\ }\textbf
  {\bibinfo {volume} {15}},\ \bibinfo {pages} {3467--3480} (\bibinfo {year}
  {2005})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {McCullough}\ \emph {et~al.}(2015)\citenamefont
  {McCullough}, \citenamefont {Small}, \citenamefont {Stemler},\ and\
  \citenamefont {Iu}}]{MichaelChaos2015}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
  {McCullough}}, \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Small}},
  \bibinfo {author} {\bibfnamefont {T.}~\bibnamefont {Stemler}}, \ and\
  \bibinfo {author} {\bibfnamefont {H.~H.-C.}\ \bibnamefont {Iu}},\ }\bibfield
  {title} {\enquote {\bibinfo {title} {Time lagged ordinal partition networks
  for capturing dynamics of continuous dynamical systems},}\ }\href {\doibase
  10.1063/1.4919075} {\bibfield  {journal} {\bibinfo  {journal} {Chaos}\
  }\textbf {\bibinfo {volume} {25}},\ \bibinfo {pages} {053101} (\bibinfo
  {year} {2015})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Kulp}\ \emph {et~al.}(2016)\citenamefont {Kulp},
  \citenamefont {Chobot}, \citenamefont {Freitas},\ and\ \citenamefont
  {Sprechini}}]{KulpChaos2016}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {C.~W.}\ \bibnamefont
  {Kulp}}, \bibinfo {author} {\bibfnamefont {J.~M.}\ \bibnamefont {Chobot}},
  \bibinfo {author} {\bibfnamefont {H.~R.}\ \bibnamefont {Freitas}}, \ and\
  \bibinfo {author} {\bibfnamefont {G.~D.}\ \bibnamefont {Sprechini}},\
  }\bibfield  {title} {\enquote {\bibinfo {title} {Using ordinal partition
  transition networks to analyze {ECG} data},}\ }\href {\doibase
  10.1063/1.4959537} {\bibfield  {journal} {\bibinfo  {journal} {Chaos}\
  }\textbf {\bibinfo {volume} {26}},\ \bibinfo {pages} {073114} (\bibinfo
  {year} {2016})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Zhang}\ \emph {et~al.}(2017)\citenamefont {Zhang},
  \citenamefont {Zhou}, \citenamefont {Tang}, \citenamefont {Guo},
  \citenamefont {Small},\ and\ \citenamefont {Zou}}]{zhangSciRep2017}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.~Y.}\ \bibnamefont
  {Zhang}}, \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont {Zhou}},
  \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Tang}}, \bibinfo {author}
  {\bibfnamefont {H.}~\bibnamefont {Guo}}, \bibinfo {author} {\bibfnamefont
  {M.}~\bibnamefont {Small}}, \ and\ \bibinfo {author} {\bibfnamefont
  {Y.}~\bibnamefont {Zou}},\ }\bibfield  {title} {\enquote {\bibinfo {title}
  {Constructing ordinal partition transition networks from multivariate time
  series},}\ }\href {\doibase 10.1038/s41598-017-08245-x} {\bibfield  {journal}
  {\bibinfo  {journal} {Scientific Reports}\ }\textbf {\bibinfo {volume} {7}},\
  \bibinfo {pages} {7795} (\bibinfo {year} {2017})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {McCullough}\ \emph {et~al.}(2017)\citenamefont
  {McCullough}, \citenamefont {Small}, \citenamefont {Iu},\ and\ \citenamefont
  {Stemler}}]{McCullough2017b}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
  {McCullough}}, \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Small}},
  \bibinfo {author} {\bibfnamefont {H.~H.~C.}\ \bibnamefont {Iu}}, \ and\
  \bibinfo {author} {\bibfnamefont {T.}~\bibnamefont {Stemler}},\ }\bibfield
  {title} {\enquote {\bibinfo {title} {{Multiscale ordinal network analysis of
  human cardiac dynamics}},}\ }\href {\doibase 10.1098/rsta.2016.0292}
  {\bibfield  {journal} {\bibinfo  {journal} {Philos. Trans. R. Soc. A Math.
  Phys. Eng. Sci.}\ }\textbf {\bibinfo {volume} {375}},\ \bibinfo {pages}
  {20160292} (\bibinfo {year} {2017})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Borges}\ \emph {et~al.}(2019)\citenamefont {Borges},
  \citenamefont {Ramos}, \citenamefont {Mini}, \citenamefont {Rosso},
  \citenamefont {Frery},\ and\ \citenamefont {Loureiro}}]{BorgesAMC2019}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.~B.}\ \bibnamefont
